\textbf{\sc Deliberative Autonomy for Marine Robotics}

### 2012-01-23 trunk/arch.tex
Short description of T-REX architecture … still not very clear

{\em This section presents quickly T-REX architecture but move
  quickly the focus on a single reactor and gross concepts on the
  architecture:
  \begin{itemize}
  \item For t-trex the whole decision and control problem is reduced
    around the state variables (or timeline construct). 
  \item This reduce the execution tracking problem for each reactor on a state
    identification problem (deduce internal state from the external
    state), and the control problem to goal posting which will trigger
    deliberation on the owner of the corresponding timeline. 
  \item the architecture itself see each reactor as a black box and
    therefore each reactor can implement its own mechanism in order to
    resolve both synchronization and deliberation. Still we provide a
    reactor based on the europa framework that leverage the automated
    planning capabilities in order to do model based planning and
    execution with a rich representation of resources and time.
  \end{itemize}}

>>>>

% {\em\tiny This section introduces shortly T-REX architecture but move
%   quickly the focus on a single reactor and gross concepts on the
%   architecture:
%   \begin{itemize}
%   \item For t-trex the whole decision and control problem is reduced
%     around the state variables (or timeline construct). 
%   \item This reduce the execution tracking problem for each reactor on a state
%     identification problem (deduce internal state from the external
%     state), and the control problem to goal posting which will trigger
%     deliberation on the owner of the corresponding timeline. 
%   \item the architecture itself see each reactor as a black box and
%     therefore each reactor can implement its own mechanism in order to
%     resolve both synchronization and deliberation. Still we provide a
%     reactor based on the europa framework that leverage the automated
%     planning capabilities in order to do model based planning and
%     execution with a rich representation of resources and time.
%   \end{itemize}}

The T-REX architecture was designed with the objective to bring 
planning closer to the low-level control of the vehicle while
maintaining its reactivity. Indeed, in most of the architecture where
task planning is embedded, this task remains at a failry high level of
abstraction as the complexity of this task would otherwise impede the
system reactiveness. As a result the planning problem managed {\em in
  situ} remains fairly detached from low-level control and most of the 
reactive adaptation which in contrast is managed by specialized 
component that relies on different representations for modeling 
the local behavior they control or even do not have any explicit 
formal model of this behavior. Taking cues from the IDEA 
architecture \cite{mus02}, we designed a core framework that provides a
formal basis \cite{Py:2010ti} on the way each of the sub-component --
called reactors -- of our agent can interact by exchanging only
information through {\em state variables} by ensuring that state 
information can propagate through all the reactors ensuring a
consistent view of the present state of the world at every single tick
trough a bottom-up {\em synchronization} flow and collaboration for
future state evolution by timely exchanges of goals on {\em state
  variables}. This exchange does not impose to any reactor to be aware
of the reactors it is connected to but only focus on the state
variables it relies on (called {\em external}) and the state
variables it manages and maintains (called {\em internal}).

Take for example the T-REX agent instance we use in our AUV
(Fig. \ref{fig:agent}). In this architecture all the reactors
(symbolized by colored boxes) do not have explicit connection from one
to another but instead rely on a publish/subscribe model on state
variables. For Example the \textsf{Pilot} reactor needs the {\em
  externally} managed state informations from \textsf{Sensor\_Data} 
and \textsf{Nav\_cmd} states variables in order to maintain {\em
  internally} the state variables \textsf{Waypoint} and
\textsf{Nav\_policy}. This relation applies in two way :
\begin{itemize}
\item In order to identify its current {\em internal} state values
  this reactor needs to know the current {\em external} state
  information it relies on. In this case the current \textsf{Waypoint}
  the vehicle is heading to can be extracted for example from the
  current \textsf{Nav\_cmd} executed.
\item The future objectives a reactor has on its future \textsf{Internal}
  state will probably imply sub-objectives to be reported to the
  owner of its {\em external} depending on their current values. Again
  if this reactor wants to go ato a specific \textsf{Waypoint} --
  depending on its current position provided by \textsf{Sensor\_Data}
  it can identify a sequence of \textsf{Nav\_cmd} that should
  eventually help him reach this location.
\end{itemize}

\begin{figure}[!htb]
 \centering
 \includegraphics[scale=0.45]{figs/AUV-agent.pdf}
 \caption{\small A \texttt{T-REX} agent is composed of multiple
   reactors or control loops (rounded boxes) which are connected
   through state variables provided by one reactor ({\color{blue}blue}
   solid line) with multiple possible clients ({\color{green}green}
   dashed lines)}
  \label{fig:agent}
\end{figure}

Overal T-REX architecture helps to abstract to the reactor interaction
to only state variables. Each reactor is practically ignoring where
its {\em external} state information is managed as long as this
information is available and properly maintained so they can properly 
manage their own {\em internal} state information which in turn may be
used by other reactors. Such formalism also constrain reactors
interaction to state information exchanges which are two folds:
\begin{itemize}
\item {\em Observation on the current state values}: This information
  propagates up in the reactor graph at the synchronization phase 
  and occur at the beginning of every time steps (called
  ticks). During this phase the agent provides to each reactor 
  updates on their {\em external} state variables for this tick and
  ask them to compute their {\em internal} state variables values for
  this same tick. By doing so we propagate a consistent view of that
  state of the world throughout all the reactors as time advance.
\item {\em Goals request on future state}: This information propagates
  down in the reactor graph. It reflects that, in order to satisfy own
  of its {\em internal} objectives, a reactor may have a plan that
  rely on future values of one of its {\em external} state. When such
  a plan is identified we ensure that this part of the plan is given
  as a goal to the owner of the given {\em external(s)} state variable(s). 
\end{itemize}

The overal execution cycle of the agent is emerging through these
interaction and the processing that occurs in each reactors in order
to process them. New observation occurring at every tick inform each
reactor of the current state of the world and impose them to identify
how it impact their state (and potentially what they planned for the
future). Goals they receive will trigger in the reactor the need to
identify a strategy (or plan) that will allow them to reach these new
states knowing their current state deduced from synchronization.When
such a plan has been identified, the {\em external}  part is then
dispatched to other reactors that can do the same. This flow and focus
around state and goals is perfectly fitting with automated planning
and even more so to the Europa framework. As an illustration, in 
 Fig. \ref{fig:agent}, only \textsf{Shore comm.} and \textsf{Vehicle}
 reactors are specialized C++ code (acting as interface to external
 programs). All the other reactors are based on the europa framework
 differentiated only by the domain model they manipulate and som
 heuristic configuration parameters. In the following sections we are
 going to focus specifically ion this kind of reactor to illustrate
 how  a formal planning framework can be used in our architecture to
 both track state evolutions and plan for future objectives while
 remaining efficient enough to do these tasks in a real-time control 
loop.
